---
title: "#DockerCon 2017"
author: "Ben Anderson (b.anderson@soton.ac.uk `@dataknut`)"
date: 'Last run at: `r Sys.time()`'
output:
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
bibliography: ~/bibliography.bib
---
```{r knitrSetUp, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # do not echo code
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig_caption = TRUE)
knitr::opts_chunk$set(fig_height = 6) # default, make it bigger to stretch vertical axis
knitr::opts_chunk$set(fig_width = 8) # full width
knitr::opts_chunk$set(tidy = TRUE) # tidy up code in case echo = TRUE
```

```{r codeSetup, include=FALSE}
# Housekeeping ----
rm(list=ls(all=TRUE)) # remove all objects from workspace

# Set start time ----
startTime <- Sys.time()
```

```{r setFunctions}
# Functions used

# This is a function to install any packages that are not present
# Especially useful when running on virtual machines where package installation is not persistent. Like UoS sve :-(
# It will fail if the packages need to be installed but there is no internet access
# Courtesy of Luke Blunden
lb_myRequiredPackages <- function(x,y){
  for( i in x ){
    #  require returns TRUE if it was able to load package
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , repos=y , 
                        #type="win.binary" , comment out so runs on OS X etc
                        quiet=TRUE , dependencies = TRUE , verbose = FALSE )
      #  Load package after installing
      require( i , character.only = TRUE, quietly = TRUE )
    }
  }
}

ba_tidyNum <- function(number) { 
  # puts commas in a long number and stops scientific format
  format(number, big.mark=",", scientific=FALSE)
}

# some things we'll use later
ba_setUseFullTimes <- function(dt){
  # set to central time
  dt <- dt[, createdLocal := with_tz(created, tz = "America/Chicago")] # Austin local
  # convert created to minutes
  dt <- dt[,
           obsDateTimeMins := floor_date(createdLocal, # requires lubridate
                                         unit="minutes"
           )
           ]
    dt <- dt[,
           obsDateTimeHours := floor_date(createdLocal, # requires lubridate
                                         unit="hours"
           )
           ]
  dt <- dt[,
           obsDateTime5m := floor_date(createdLocal, # requires lubridate
                                         unit="5 minutes"
           )
           ]
  dt <- dt[,
           obsDateTime10m := floor_date(createdLocal, # requires lubridate
                                         unit="10 minutes"
           )
           ] 
  dt <- dt[,
           obsDateTime15m := floor_date(createdLocal, # requires lubridate
                                         unit="15 minutes"
           )
           ] 
  
  dt <- dt[, obsDate := as.Date(obsDateTimeMins, tz = "America/Chicago")]

  # test
  #t <- tweetListDT[, .(createdLocal, obsDate, obsDateTimeMins)]
  
  dt <- dt[, isRetweetLab := ifelse(isRetweet == "FALSE",
                                    "Original tweet",
                                    "Re-tweet"
                                    )
           ]
  # check
  #table(tweetListDT$isRetweet,tweetListDT$isRetweetLab)
  return(dt)
}

ba_make5MinTimeSeriesChart <- function(dt,byVars,facetForm){
  # obsDateTime5m must be one of the byVars
  # whatever is in facetForm must also be in byVars
  plotDT <- dt[,
                 .(
                   nTweets = .N,
                   nTweeters = uniqueN(screenName)
                 ), by = eval(byVars)
               ]
  

  myPlot <- ggplot(plotDT, aes(x = obsDateTime5m)) +
    geom_line(aes(y = nTweets, colour = "N tweets")) +
    geom_line(aes(y = nTweeters, colour = "N tweeters")) +
    facet_grid(eval(facetForm)) +
    theme(strip.text.y = element_text(size = 9, colour = "black", angle = 90)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
    scale_x_datetime(date_breaks = "2 hours", date_labels ="%a %d %b %H:%M") +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) +
    labs(caption = myCaption,
         x = "Time",
         y = "Count"
    )
  return(myPlot)
}
```

```{r setPackages}
# Load libraries ----
# Use the function to load the libraries required by this script
reqLibs <- c("data.table", # fast data munching
             "lubridate", # for date manipulation
             "ggplot2", # fancy graphs
             "readr", # fast data loading e.g. read_csv
             "plotly", # scalable graphs in .js
             "knitr" # for kable
)

print(paste0("Loading the following libraries using lb_myRequiredPackages: ", reqLibs))
# Use Luke's function to require/install/load
lb_myRequiredPackages(reqLibs,"http://cran.rstudio.com/")

# uncomment to install twitteR if needed
#library(devtools)
#install_github("twitteR", username="geoffjentry", force = TRUE)
library(twitteR)
```


# Purpose

To extract and visualise tweets and re-tweets of `#dockercon` for 17 - 21 April, 2017 (DockerCon17).

Borrowing extensively from http://thinktostart.com/twitter-authentification-with-r/

We used the Twitter search API to extract 'all' tweets with the `#dockercon` hashtag. As the [Twitter search API documentation](https://dev.twitter.com/rest/public/search) (sort of) makes clear this may not be `all` such tweets but merely the `most relevant` (whatever that means) from within a `sample` (whatever that means). 

>"It allows queries against the indices of recent or popular Tweets and behaves similarly to, but not exactly like the Search feature available in Twitter mobile or web clients, such as Twitter.com search. The Twitter Search API searches against a sampling of recent Tweets published in the past 7 days." https://dev.twitter.com/rest/public/search, Accessed 12/5/2017

It is therefore possible that not quite all tweets have been extracted although it seems likely that we have captured most `human` tweeting which was our main intention. Future work should instead use the Twitter [streaming API](https://dev.twitter.com/streaming/overview).

# Load Data
Data should have been already downloaded using `collectData.R`. This produces a data table with the following variables (after some processing):

```{r loadData}
tweetListDT <- data.table()

# load the 6 days of data (due to twitter search being in UTC)
for(d in 1:6){
  file <- paste0("tweetListD", d, "DT.csv")
  tweetListDT <- rbind(tweetListDT,data.table(read_csv(file)))
}

#tweetListDT <- data.table(read_csv("tweetListDT.csv")) # requires readr, auto-parses dateTimes nicely

# add stuff (especially dates & times corrected to local Austin time)
tweetListDT <- ba_setUseFullTimes(tweetListDT)

# keep the data within the time period of interest
# note that this selects according to CDT which seems the most relevant
tweetListDT <- tweetListDT[as.Date(obsDateTimeMins) >= as.Date("2017-04-17", tz = "America/Chicago") &
                             as.Date(obsDateTimeMins) <= as.Date("2017-04-21", tz = "America/Chicago")
                             ]



names(tweetListDT)
```

The table has `r ba_tidyNum(nrow(tweetListDT[isRetweet == "FALSE"]))` tweets (and `r ba_tidyNum(nrow(tweetListDT[isRetweet == "TRUE"]))` re-tweets) from `r ba_tidyNum(uniqueN(tweetListDT$screenName))` tweeters between `r min(tweetListDT$createdLocal)` and `r max(tweetListDT$createdLocal)` (Central District Time).

# Analysis

## Tweets and Tweeters over time

```{r setCaptionTimeSeries}
myCaption <- paste0("All (re)tweets containing #dockercon ", 
                      min(as.Date(tweetListDT$obsDateTime5m)),
                          " to ",
                          max(as.Date(tweetListDT$obsDateTime5m))
                          )
```

```{r allDaysChart, fig.height=8, fig.width=9, fig.cap=myCaption}

myPlot <- ba_make5MinTimeSeriesChart(tweetListDT, 
                                byVars = c("obsDateTime5m,isRetweetLab"),
                                facetForm = c("isRetweetLab ~ .")
                                )

myPlot

#ggplotly(myPlot)
```

### Day 1 - Monday (Workshops)

This plot is zoomable - try it!

```{r day1Chart, fig.height=8, fig.width=9, fig.cap=myCaption}
myCaption <- paste0("All (re)tweets containing #dockercon Monday 17th April 2017")

myPlot <- ba_make5MinTimeSeriesChart(tweetListDT[obsDate == "2017-04-17"], 
                                byVars = c("obsDateTime5m,isRetweetLab"),
                                facetForm = c("isRetweetLab ~ .")
                                )

#myPlot

ggplotly(myPlot)
```

### Day 2 - Tuesday (Main Day 1)

This plot is zoomable - try it!

```{r day2Chart, fig.height=8, fig.width=9, fig.cap=myCaption}
myCaption <- paste0("All (re)tweets containing #dockercon Tuesday 18th April 2017")

myPlot <- ba_make5MinTimeSeriesChart(tweetListDT[obsDate == "2017-04-18"], 
                                byVars = c("obsDateTime5m,isRetweetLab"),
                                facetForm = c("isRetweetLab ~ .")
                                )
myPlot <- myPlot + 
  annotate("rect", 
           xmin = as.POSIXct("2017-04-18 09:00:00"), 
           xmax = as.POSIXct("2017-04-18 09:15:00"), 
           ymin = 0, 
           ymax = 100, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-18 09:00:00"), 
           y = 100,
           label = c("Ben Golub: Day 1 keynote")
           ) 


ggplotly(myPlot)
```


### Day 3 - Wednesday (Main Day 2)

This plot is zoomable - try it!

```{r day3Chart, fig.height=8, fig.width=9, fig.cap=myCaption}
myCaption <- paste0("All (re)tweets containing #dockercon Wednesday 19th April 2017")
myPlot <- ba_make5MinTimeSeriesChart(tweetListDT[obsDate == "2017-04-19"], 
                                byVars = c("obsDateTime5m,isRetweetLab"),
                                facetForm = c("isRetweetLab ~ .")
                                )
myPlot <- myPlot + annotate("rect", 
           xmin = as.POSIXct("2017-04-19 09:00:00"), 
           xmax = as.POSIXct("2017-04-19 10:30:00"), 
           ymin = 0, 
           ymax = 80, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-19 09:00:00"), 
           y = 90,
           label = c("Day 2: General session 1")
           ) +
  annotate("rect", 
           xmin = as.POSIXct("2017-04-19 16:45:00"), 
           xmax = as.POSIXct("2017-04-19 17:30:00"), 
           ymin = 0, 
           ymax = 80, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-19 16:45:00"), 
           y = 90,
           label = c("Day 2: Cool Hacks")
           )
# annotate("rect", 
#            xmin = as.POSIXct("2017-04-19 16:45:00"), 
#            xmax = as.POSIXct("2017-04-19 16:50:00"), 
#            ymin = 0, 
#            ymax = 100, alpha = .2) +
#   annotate("text", 
#            x = as.POSIXct("2017-04-19 16:45:00"), 
#            y = 100,
#            label = c("Docker Cool hacks: \n@xetorthio @marcosnils")
#            ) +
# annotate("rect", 
#            xmin = as.POSIXct("2017-04-19 17:05:00"), 
#            xmax = as.POSIXct("2017-04-19 17:10:00"), 
#            ymin = 0, 
#            ymax = 100, alpha = .2) +
#   annotate("text", 
#            x = as.POSIXct("2017-04-19 17:05:00"), 
#            y = 80,
#            label = c("Docker Cool hacks: \n@alexellisuk")
#            ) +
# annotate("rect", 
#            xmin = as.POSIXct("2017-04-19 17:20:00"), 
#            xmax = as.POSIXct("2017-04-19 17:25:00"), 
#            ymin = 0, 
#            ymax = 100, alpha = .2) +
#   annotate("text", 
#            x = as.POSIXct("2017-04-19 17:20:00"), 
#            y = 60,
#            label = c("Ben Golub: \nDockerCon18")
#            )


ggplotly(myPlot)
```


### Day 4 - Thursday (Main Day 3)

```{r day4Chart, fig.height=8, fig.width=9, fig.cap=myCaption}
myCaption <- paste0("All (re)tweets containing #dockercon Thursday 20th April 2017")
myPlot <- ba_make5MinTimeSeriesChart(tweetListDT[obsDate == "2017-04-20"], 
                                byVars = c("obsDateTime5m,isRetweetLab"),
                                facetForm = c("isRetweetLab ~ .")
                                )
myPlot


#ggplotly(myPlot)
```


## Location (lat/long)
We wanted to make a nice map but sadly we see that most tweets have no lat/long set.

```{r latLongPlot}
plotDT <- tweetListDT[, 
                    .(
                      nTweets = .N
                    ), by = .(latitude, longitude)]
kable(cap="All logged lat/long values",
      plotDT)
```

## Location (textual)

This appears to be pulled from the user's profile although it may also be a 'guestimate' of current location.

Top locations for tweets:

```{r tweetLocations}
t <- tweetListDT[, .(nTweets = .N), by = location][order(-nTweets)]

kable(caption = "Top 15 locations for tweeting",
      head(t, 15)
      )
```

Top locations for tweeters:

```{r tweeterLocations}
t <- tweetListDT[, .(nTweeters = uniqueN(screenName)), by = location][order(-nTweeters)]

kable(caption = "Top 15 locations for tweeters",
      head(t, 15)
      )
```


## Screen name

Next we'll try by screen name.

Top tweeters:

```{r topTweeters}
allTweetersDT <- tweetListDT[, .(nTweets = .N), by = screenName][order(-nTweets)]

kable(caption = "Top 15 tweeters",
      head(allTweetersDT, 15)
      )
```

And here's a really bad visualisation of all of them tweeting over time! Each row of pixels is a tweeter (the names are illegible) and a green dot indicates a few tweets in the 5 minute period while a red dot indicates a lot of tweets.

```{r screenNameAll, fig.height=8,fig.cap="N tweets per 5 minutes by screen name"}
myCaption <- paste0("All (re)tweets containing #dockercon ", 
                      min(as.Date(tweetListDT$obsDateTime5m)),
                          " to ",
                          max(as.Date(tweetListDT$obsDateTime5m))
                          )

plotDT <- tweetListDT[, 
                    .(
                      nTweets = .N
                    ), by = .(screenName, obsDateTime5m)]

myPlot <- ggplot(plotDT, aes(x = obsDateTime5m)) +
    geom_tile(aes(y = screenName, fill = nTweets)) +
    theme(strip.text.y = element_text(size = 9, colour = "black", angle = 0)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
    #scale_x_reverse() + # fix reverse plotting of long
    scale_x_datetime(date_breaks = "2 hours", date_labels ="%d %b %H:%M") +
    scale_fill_gradient(low="green", high = "red") +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) +
    labs(caption = myCaption,
         x = "Time",
         y = "Screen name"
    )

#ggplotly(myPlot)
ymax <- uniqueN(plotDT$screenName)
myPlot + annotate("rect", 
           xmin = as.POSIXct("2017-04-18 09:00:00"), 
           xmax = as.POSIXct("2017-04-18 09:15:00"), 
           ymin = 0, 
           ymax = ymax, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-18 09:00:00"), 
           y = ymax - 10,
           label = c("Ben Golub: Day 1 keynote")
           ) +
  annotate("rect", 
           xmin = as.POSIXct("2017-04-19 09:00:00"), 
           xmax = as.POSIXct("2017-04-19 10:30:00"), 
           ymin = 0, 
           ymax = ymax, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-19 09:00:00"), 
           y = ymax-10,
           label = c("Day 2: General session 1")
           ) +
  annotate("rect", 
           xmin = as.POSIXct("2017-04-19 16:45:00"), 
           xmax = as.POSIXct("2017-04-19 17:30:00"), 
           ymin = 0, 
           ymax = ymax, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-19 16:45:00"), 
           y = ymax-10,
           label = c("Day 2: Cool Hacks")
           )
```

So let's re-do that for the top 50 tweeters so we can see their tweetStreaks!

```{r screenNameTop50, fig.height=8,fig.cap="N tweets per 5 minutes by screen name (top 50, most prolific tweeters at bottom)"}
myCaption <- paste0("All (re)tweets containing #dockercon ", 
                      min(as.Date(tweetListDT$obsDateTime5m)),
                          " to ",
                          max(as.Date(tweetListDT$obsDateTime5m)),
                    "\nOrdered by most frequent tweeters (bottom)"
                          )

matchDT <- head(allTweetersDT,50)
matchDT <- matchDT[, maxT := nTweets]
setkey(matchDT, screenName)
setkey(tweetListDT, screenName)

tempDT <- merge(tweetListDT, matchDT)

plotDT <- tempDT[matchDT, 
                    .(
                      nTweets = .N
                    ), by = .(screenName, obsDateTime5m, maxT)]

plotDT <- plotDT[order(-plotDT$maxT,plotDT$obsDateTime5m)]
#plotDT$screenName <- factor(plotDT$screenName,levels=plotDT$screenName)

myPlot <- ggplot(plotDT, aes(x = obsDateTime5m)) +
    geom_tile(aes(y = screenName, fill = nTweets)) +
    theme(strip.text.y = element_text(size = 9, colour = "black", angle = 0)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
    #scale_x_reverse() + # fix reverse plotting of long
    scale_x_datetime(date_breaks = "2 hours", date_labels ="%d %b %H:%M") +
    scale_fill_gradient(low="green", high = "red") +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) +
    labs(caption = myCaption,
         x = "Time",
         y = "Screen name"
    )

#ggplotly(myPlot)
ymax <- uniqueN(plotDT$screenName)
myPlot + annotate("rect", 
           xmin = as.POSIXct("2017-04-18 09:00:00"), 
           xmax = as.POSIXct("2017-04-18 09:15:00"), 
           ymin = 0, 
           ymax = ymax, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-18 09:00:00"), 
           y = ymax - 1,
           label = c("Ben Golub: Day 1 keynote")
           ) +
  annotate("rect", 
           xmin = as.POSIXct("2017-04-19 09:00:00"), 
           xmax = as.POSIXct("2017-04-19 10:30:00"), 
           ymin = 0, 
           ymax = ymax, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-19 09:00:00"), 
           y = ymax-1,
           label = c("Day 2: General session 1")
           ) +
  annotate("rect", 
           xmin = as.POSIXct("2017-04-19 16:45:00"), 
           xmax = as.POSIXct("2017-04-19 17:30:00"), 
           ymin = 0, 
           ymax = ymax, alpha = .2) +
  annotate("text", 
           x = as.POSIXct("2017-04-19 16:45:00"), 
           y = ymax-1,
           label = c("Day 2: Cool Hacks")
           )
```

# About

Analysis completed in: `r round(Sys.time() - startTime, 2)` seconds using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

A special mention must go to `twitteR` [@twitteR] for the twitter API interaction functions and `lubridate` [@lubridate] which allows timezone manipulation without too many tears.

Other R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * readr - for nice data loading [@readr]
 * ggplot2 - for slick graphs [@ggplot2]
 * plotly - fancy, zoomable slick graphs [@plotly]
 * knitr - to create this document [@knitr]

# References


