---
title: "#DockerCon 2017"
author: "Ben Anderson (b.anderson@soton.ac.uk `@dataknut`)"
date: 'Last run at: `r Sys.time()`'
output:
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
bibliography: ~/bibliography.bib
---
```{r knitrSetUp, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # do not echo code
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig_caption = TRUE)
knitr::opts_chunk$set(fig_height = 6) # default, make it bigger to stretch vertical axis
knitr::opts_chunk$set(fig_width = 8) # full width
knitr::opts_chunk$set(tidy = TRUE) # tidy up code in case echo = TRUE
```

```{r codeSetup, include=FALSE}
# Housekeeping ----
rm(list=ls(all=TRUE)) # remove all objects from workspace

# Set start time ----
startTime <- Sys.time()
```

```{r setFunctions}
# Functions used across all SAVE code

# This is a function to install any packages that are not present
# Especially useful when running on virtual machines where package installation is not persistent. Like UoS sve :-(
# It will fail if the packages need to be installed but there is no internet access
# Courtesy of Luke Blunden
lb_myRequiredPackages <- function(x,y){
  for( i in x ){
    #  require returns TRUE if it was able to load package
    if( ! require( i , character.only = TRUE ) ){
      #  If package was not able to be loaded then re-install
      install.packages( i , repos=y , 
                        #type="win.binary" , comment out so runs on OS X etc
                        quiet=TRUE , dependencies = TRUE , verbose = FALSE )
      #  Load package after installing
      require( i , character.only = TRUE, quietly = TRUE )
    }
  }
}

ba_tidyNum <- function(number) { 
  # puts commas in a long number and stops scientific format
  format(number, big.mark=",", scientific=FALSE)
}

# some things we'll use later
ba_setUseFullTimes <- function(dt){
  # set to central time
  dt <- dt[, createdLocal := with_tz(created, tz = Sys.timezone())]
  # convert created to minutes
  dt <- dt[,
           obsDateTimeMins := floor_date(createdLocal, # requires lubridate
                                         unit="minutes"
           )
           ]
    dt <- dt[,
           obsDateTimeHours := floor_date(createdLocal, # requires lubridate
                                         unit="hours"
           )
           ]
  dt <- dt[, obsMin := as.POSIXlt(obsDateTimeMins)$min]
  dt <- dt[, obsHour := as.POSIXlt(obsDateTimeMins)$hour]
  
  dt <- dt[, obsQH := ifelse(obsMin < 15, "00", NA)]
  dt <- dt[, obsQH := ifelse(obsMin >= 15, "15", obsQH)]
  dt <- dt[, obsQH := ifelse(obsMin >= 30, "30", obsQH)]
  dt <- dt[, obsQH := ifelse(obsMin >= 45, "45", obsQH)]
  
  # set q hour (useful for charts)
  dt <- dt[, obsQHour := paste0(dt$obsHour,":", dt$obsQH)
           ]
  dt <- dt[,
           obsDateTime5m := floor_date(createdLocal, # requires lubridate
                                         unit="5 minutes"
           )
           ]
  dt <- dt[,
           obsDateTime10m := floor_date(createdLocal, # requires lubridate
                                         unit="10 minutes"
           )
           ] 
  dt <- dt[,
           obsDateTime15m := floor_date(createdLocal, # requires lubridate
                                         unit="15 minutes"
           )
           ] 
  return(dt)
}

ba_make5MinTimeSeriesChart <- function(dt,byVars,facetForm){
  # obsDateTime5m must be one of the byVars
  # whatever is in facetForm must also be in byVars
  plotDT <- dt[,
                 .(
                   nTweets = .N,
                   nTweeters = uniqueN(screenName)
                 ), by = eval(byVars)
               ]
  

  myPlot <- ggplot(plotDT, aes(x = obsDateTime5m)) +
    geom_line(aes(y = nTweets, colour = "N tweets")) +
    geom_line(aes(y = nTweeters, colour = "N tweeters")) +
    facet_grid(eval(facetForm)) +
    theme(strip.text.y = element_text(size = 9, colour = "black", angle = 0)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
    scale_x_datetime(date_breaks = "2 hours", date_labels ="%a %d %b %H:%M") +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) +
    labs(caption = myCaption,
         x = "Time",
         y = "Count"
    )
  return(myPlot)
}
```

```{r setPackages}
# Load libraries ----
# Use the function to load the libraries required by this script
reqLibs <- c("data.table", # fast data munching
             "lubridate", # for date manipulation
             "ggplot2", # fancy graphs
             "readr", # fast data loading e.g. read_csv
             "plotly", # scalable graphs in .js
             "knitr" # for kable
)

print(paste0("Loading the following libraries using lb_myRequiredPackages: ", reqLibs))
# Use Luke's function to require/install/load
lb_myRequiredPackages(reqLibs,"http://cran.rstudio.com/")

# uncomment to install twitteR if needed
#library(devtools)
#install_github("twitteR", username="geoffjentry", force = TRUE)
library(twitteR)
```


# Purpose

To extract and visualise tweets and re-tweets of `#dockercon` for 17 - 21 April, 2017 (DockerCon17).

Borrowing extensively from http://thinktostart.com/twitter-authentification-with-r/

# Load Data
Data should have been already downloaded using `collectData.R`. This produces a data table with the following variables (after some processing):

```{r loadData}
tweetListDT <- data.table(read_csv("tweetListDT.csv")) # requires readr, auto-parses dateTimes nicely

# add stuff
tweetListDT <- ba_setUseFullTimes(tweetListDT)

names(tweetListDT)
```

The table has `r ba_tidyNum(nrow(tweetListDT[isRetweet == "FALSE"]))` tweets (and `r ba_tidyNum(nrow(tweetListDT[isRetweet == "TRUE"]))` re-tweets) from `r ba_tidyNum(uniqueN(tweetListDT$screenName))` tweeters between `r min(tweetListDT$createdLocal)` and `r max(tweetListDT$createdLocal)` (Central District Time).

# Analysis

## Tweets and Tweeters over time

```{r setCaptionTimeSeries}
myCaption <- paste0("All (re)tweets containing #dockercon ", 
                      min(as.Date(tweetListDT$obsDateTime5m)),
                          " to ",
                          max(as.Date(tweetListDT$obsDateTime5m)),
                    "\nFALSE = tweets, TRUE = re-tweets"
                          )
```

```{r allDaysChart, fig.height=8, fig.width=9, fig.cap=myCaption}

ggplotly(ba_make5MinTimeSeriesChart(tweetListDT, 
                                byVars = c("obsDateTime5m,isRetweet"),
                                facetForm = c("isRetweet ~ .")
                                )
)

```

## Location (lat/long)
We want to make a nice map but sadly we see that most tweets have no lat/long set.

```{r latLongPlot}
plotDT <- tweetListDT[, 
                    .(
                      nTweets = .N
                    ), by = .(latitude, longitude)]
kable(cap="All logged lat/long values",
      plotDT)
```

## Location (textual)

This appears to be pulled from the user's profile although it may also be a 'guestimate' of current location.

Top locations for tweets:

```{r tweetLocations}
t <- tweetListDT[, .(nTweets = .N), by = location][order(-nTweets)]

kable(caption = "Top 15 locations for tweeting",
      head(t, 15)
      )
```

Top locations for tweeters:

```{r tweeterLocations}
t <- tweetListDT[, .(nTweeters = uniqueN(screenName)), by = location][order(-nTweeters)]

kable(caption = "Top 15 locations for tweeters",
      head(t, 15)
      )
```


## Screen name

Next we'll try by screen name.

Top tweeters:

```{r topTweeters}
allTweetersDT <- tweetListDT[, .(nTweets = .N), by = screenName][order(-nTweets)]

kable(caption = "Top 15 tweeters",
      head(allTweetersDT, 15)
      )
```

And here's a really bad visualisation of all of them!

```{r screenNameAll, fig.height=8,fig.cap="N tweets per 5 minutes by screen name"}
myCaption <- paste0("All (re)tweets containing #dockercon ", 
                      min(as.Date(tweetListDT$obsDateTime5m)),
                          " to ",
                          max(as.Date(tweetListDT$obsDateTime5m))
                          )

plotDT <- tweetListDT[, 
                    .(
                      nTweets = .N
                    ), by = .(screenName, obsDateTime5m)]

myPlot <- ggplot(plotDT, aes(x = obsDateTime5m)) +
    geom_tile(aes(y = screenName, fill = nTweets)) +
    theme(strip.text.y = element_text(size = 9, colour = "black", angle = 0)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
    #scale_x_reverse() + # fix reverse plotting of long
    scale_x_datetime(date_breaks = "2 hours", date_labels ="%d %b %H:%M") +
    scale_fill_gradient(low="green", high = "red") +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) +
    labs(caption = myCaption,
         x = "Time",
         y = "Screen name"
    )

myPlot
```

So let's re-do that for the top 50 tweeters.

```{r screenNameTop50, fig.height=8,fig.cap="N tweets per 5 minutes by screen name (top 50)"}
myCaption <- paste0("All (re)tweets containing #dockercon ", 
                      min(as.Date(tweetListDT$obsDateTime5m)),
                          " to ",
                          max(as.Date(tweetListDT$obsDateTime5m)),
                    "\nOrdered by most frequent tweeters (bottom)"
                          )

matchDT <- head(allTweetersDT,50)
matchDT <- matchDT[, maxT := nTweets]
setkey(matchDT, screenName)
setkey(tweetListDT, screenName)

tempDT <- merge(tweetListDT, matchDT)

plotDT <- tempDT[matchDT, 
                    .(
                      nTweets = .N
                    ), by = .(screenName, obsDateTime5m, maxT)]

plotDT <- plotDT[order(-plotDT$maxT,plotDT$obsDateTime5m)]
plotDT$screenName <- factor(plotDT$screenName,levels=plotDT$screenName)

myPlot <- ggplot(plotDT, aes(x = obsDateTime5m)) +
    geom_tile(aes(y = screenName, fill = nTweets)) +
    theme(strip.text.y = element_text(size = 9, colour = "black", angle = 0)) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
    #scale_x_reverse() + # fix reverse plotting of long
    scale_x_datetime(date_breaks = "2 hours", date_labels ="%d %b %H:%M") +
    scale_fill_gradient(low="green", high = "red") +
    theme(legend.position = "bottom") +
    theme(legend.title = element_blank()) +
    labs(caption = myCaption,
         x = "Time",
         y = "Screen name"
    )

myPlot
```

# About

Analysis completed in: `r round(Sys.time() - startTime, 2)` seconds using [knitr](https://cran.r-project.org/package=knitr) in [RStudio](http://www.rstudio.com) with `r R.version.string` running on `r R.version$platform`.

A special mention must go to [`twitteR`](https://github.com/geoffjentry/twitteR) for the twitter API interaction functions and `lubridate` [`lubridate`](https://github.com/hadley/lubridate) which allows timezone manipulation without tears.

Other R packages used:

 * base R - for the basics [@baseR]
 * data.table - for fast (big) data handling [@data.table]
 * readr - for nice data loaading [@readr]
 * ggplot2 - for slick graphs [@ggplot2]
 * plotly - fancy, zoomable slick graphs [@plotly]
 * knitr - to create this document [@knitr]

# References


